---
title: "Untitled"
author: "Fomba Kassoh"
date: "2023-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(readr)

spotify_data <- read_csv('C:/Users/RemoteUser/Documents/DATA 607/Tidyverse Create/spotify_songs.csv')
glimpse(spotify_data)
```

ChatGPT
A box plot, also known as a box-and-whisker plot, is a standardized way of displaying the distribution of data based on a five-number summary: minimum, first quartile (Q1), median, third quartile (Q3), and maximum. It can tell you about your outliers and what their values are. It can also tell you if your data is symmetrical, how tightly your data is grouped, and if and how your data is skewed.

Here’s how you can use a box plot for data analysis:

Central Tendency: The line in the middle of the box indicates the median, which is a good indicator of the central tendency of the data.

Spread: The size of the box shows the interquartile range (IQR), which is the distance between the first and third quartiles. It indicates the spread of the middle half of the data. Smaller boxes mean less variability in the data.

Skewness: If the median is closer to the bottom of the box and the whiskers are unequal lengths, the data is positively skewed. If the median is closer to the top of the box, the data is negatively skewed.

Outliers: Points plotted beyond the whiskers are outliers and are often worth further investigation. They can indicate variability outside the common statistics or issues with data collection.

Comparison: When you have multiple box plots side by side (for example, comparing different groups), it's easier to compare medians, spreads, and the range of the data across these groups.

```{r}
library(ggplot2)

# Assuming your data is in a data frame named 'df'
p <- ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity)) + 
  geom_boxplot() +
  theme_minimal() + 
  labs(title = "Distribution of Track Popularity by Playlist Genre",
       x = "Playlist Genre",
       y = "Track Popularity")

print(p)

```




